---
title: "Analytics for Activity Trackers: Project Proposal"
output:
     html_document
---

__Proposal Deadline: Jan 30, 2017 11:59pm__


- Edit this markdown file by __replacing only__ what is there in the
chunks named `q0`, `q1`, `q2`, ..., `q5` with your team's appropriate
answers. These chunks have been filled in as an example.

- Place the filled in markdown in any one of the team member's private
directory with the same name `proposal.Rmd`. Do not change the name!


0. Who are the authors of this package? See format below.

```{r q0}
authors <- list(radke = "Rohisha Adke",
                layu = "Lisa Ann Yu")
```

1. What is your package trying to achieve?  Who is it aimed at?

```q1
1. Data matching. The package will allow users to input data from their personal 
activity trackers or other logs, and match the data sources together for a 
multifaceted view of their health. The primary personal tracking data the 
package will be used with is Fitbit data, which we will access for a user 
through the R package fitbitScraper. Users may also input their own sources of 
data as files, such as Apple Health or other app data.

2. Visualization. The package will allow users to visualize their matched data. 
To increase the ease of use for Fitbit users, these will also be displayed in a 
Shiny app. For example,

  - Comparisons between Fitbit data and Apple Health data for matched fields.
  Note: the visualizations, analyses, recommendations, and framework will all 
  rely on Fitbit data, since it collects more data than Apple Health
  - Multiple measures (ex: for Fitbit data, heartrate, step count/distance, 
  calories burned, weight) over the course of a day (or any other time frame), 
  on the same graph (normalized, or with multiple axes).
  - For workouts, can plot rates (distance/time over time) or distances - useful 
  if someone is trying to improve their mile time or prep for an event (5k, 
  marathon)
  - User can plug in other values, like when the workday starts and ends,
  and when they had caffeine or alcohol, to compare all other values to those
  time points.
  - Compare to average in gender/age bracket, compare to own data at an earlier
  time (last week, last month, etc.).
  
3. Analysis. See associations between data types - for example, the relationship 
between steps and distance during workouts, comparing sleep pattern to daytime 
data, comparing weekend and weekday data, etc.

4. Recommendations for improvement. Based on a user's choice of goal, we can 
recommend strategies based on their data (for example, if a user walks a lot on 
the weekdays but not weekends and is not meeting their target step count, 
suggest walks on the weekends, at times of day when they are usually active on 
the weekdays according to their heart rate).
  
5. Experimentation framework. The package will provide the framework for a user 
to easily perform A/B or other testing on him/herself - for example, decide 
he/she wants to know if his/her heart rate is higher in a statistically 
significant way when alcohol or caffeine is consumed, and so mark each day/time 
of consumption, then display the results of hypothesis testing on the desired field.


The package is aimed at users of Fitbits and other self-tracking devices and methods looking to get more interaction and insight from their data.

```

2. Briefly describe the background if that's not obvious.  What field
   or area of interest is it related to?  (Just briefly, in a paragraph
   or two, maximum.) References, including web references, are welcome
   where available.

```q2
Fitbits and other health tracking devices have taken off in the last few years, 
as data enthusiasts, along with, increasingly, the everyday consumer, seek a 
more personal, and finer, granularity of insight. There are stories of devices 
alerting users to previously unknown health conditions, while at the same time, 
how to incorporate such data into medical care remains an open question.

A common complaint about the Fitbit in particular is the lack of analytics 
provided for by the company; users are left to glance through a couple raw 
graphs of their numbers, and, failing to continue to learn from their data, 
after the novelty wears off, Fitbits often end up stuffed away, unused, in a 
drawer.

This package will bring the Fitbit one step closer to its full potential by 
providing users with suggestions and the ability to do advanced analysis of 
their health. Instead of single graph visualizations of raw numbers, we will 
enable individuals to view the impact of one part of their life on another by 
combining data sources. We hope to allow the Fitbit to remain an interactive 
tool for self analysis and tailoring of one's life to meet one's goals, and a 
source of continual learning, by enabling individuals to run experiments in 
their life, then easily view the impact of the treatment on their Fitbit data.


Related work includes:

The fitbitScraper R package, which allows users to access their Fitbit data: https://cran.r-project.org/web/packages/fitbitScraper/index.html.
The analytics requests of the Fitbit community: https://community.fitbit.com/t5/Fitbit-com-Dashboard/dashboard-needs-better-analytics/m-p/1185213
Other websites that create Fitbit analytics: https://rocketgraph.com/reports/16-fitbit-analytics, https://www.producthunt.com/posts/fitbit-analytics-2, https://datarunsdeep.com.au/blog/track-your-life-in-google-analytics/
```

3. Provide one or two use cases; that is, in a sentence or so say how
   you see someone using the package.

```q3
Individuals can use the R package to match data from their personal health 
devices and apps together for global analysis.

Individuals can use the R package and Shiny app to run experiments on their 
Fitbit data; for example, track when they consume alcohol, and visualize the 
result on their recorded Fitbit resting heartrate.

Individuals can run analytics on their data, such as an analysis of how long 
they usually spend between waking up, and leaving bed in the morning.

Individuals can visualize associations in their data, such as the result of 
exercise and increased step counts on their sleep at night.

Individuals can view fluctuations in their measures of health over time, such as 
how their heart rate corresponds to other events as recorded by other tracking 
devices.

Individuals can design a training program, or use a suggested one, and measure 
the effect on their weight over time.
```

4. If you can, list the main functions or other software you plan on
   including.  No need to implement anything now, but you should have
   an idea.

```q4
The R package will provide capability to read in Fitbit, Apple, and other 
personal health data, match data from multiple sources, visualize the data and 
associations between different types of data, perform analysis on the data, and 
run experiments over time by adding other types of data to their personal tracking data.
```

```{r, echo = FALSE, hide = TRUE}
none <- "none"
basic <- "basic"
proficient <- "proficient"
```

5. For languages C/C++, Java, Python, Ruby, please specify your
   proficiency: Proficient, basic knowledge, no knowledge.  (This
   helps us evaluate the proposal sometimes and offer suggestions.)

```{r q5}
languages <- list(
    ## replace the values below appropriately for each member
    radke = c(cxx = none,
               java = basic,
               python = proficient,
               ruby = none)
    ,
    layu = c(cxx = none,
               java = basic,
               python = basic,
               ruby = none)
)
```

