---
title: "Analytics for Activity Trackers: Project Proposal"
output:
     html_document
---

__Proposal Deadline: Jan 30, 2017 11:59pm__


- Edit this markdown file by __replacing only__ what is there in the
chunks named `q0`, `q1`, `q2`, ..., `q5` with your team's appropriate
answers. These chunks have been filled in as an example.

- Place the filled in markdown in any one of the team member's private
directory with the same name `proposal.Rmd`. Do not change the name!


0. Who are the authors of this package? See format below.

```{r q0}
authors <- list(radke = "Rohisha Adke",
                layu = "Lisa Ann Yu")
```

1. What is your package trying to achieve?  Who is it aimed at?

```q1

'''
1.  Querying.  The package will enable users to query the Adverse
Events Database and other databases on OpenFDA from R. For example,
they could query information about events and patients.

2.  Visualization.  The package will also provide tools to
visualize for an inquired drug the relationship between adverse event
counts/types and time, drug pair, patient's weight, age, gender and
other variables.

3.  Analysis.  It will also include three functions to analyze
if 1) there are spikes in adverse event happenings, 2) if spikes
are related to drug recalls or enforcement and then 3) other
visualizations such as principal component charts to see which
variables contribute the most to certain phenomena.

The package is aimed at analysts and scientists in pharmaceutical
companies, research institutions, health care organizations, and other
administrative organizations (i.e., FDA).
'''

Ideas **just have to make sure doing more than this package we are using does,
including the examples in the vignette**:
- Visualization: 
  - across the course of a day (or any other time frame), several different
  measures (i.e. heart-rate, steps, distance, calories burned) on one graph
  so the user can look at the relationship between those variables; can have
  two different variables on a plot on two different scales, or several on one
  plot just using densities for all the variables
  - for workouts, can plot rates (distance/time over time) or distances - useful 
  if someone is trying to improve their mile time or prep for an event (5k, 
  marathon)
  - the user can plug in other values, like when the workday starts and ends,
  and when they had caffeine or alcohol, to compare all other values to those
  time points
  - compare to average in gender/age bracket
  (I guess we will have to scrape this?), compare to own data at an earlier
  time (last week, last month, etc.)
- Recommendations for improvement: 
  - this will be hard: we can show the user is not meeting his/her own goals 
  but how can we write a program that will provide nonobvious suggestions?
- Analysis:
  - what counts as an analysis? we could look at the relationship between
  steps and distance during "workout sessions" and do linear regressions?
  - compare sleep to what happened during the day - use lag and lead
- Shiny App: users will be able to use the app to look at these graphs/analytics 
themselves



1. Data matching. The package will allow users to input data from their personal activity trackers or other logs, and match the data sources together for global analysis of their health. The primary personal tracking data the package will be used to deal with is Fitbit data, which we will use the R package fitbitScraper to access. However, users may also input their own alternative sources of data as files, such as Apple Health data, or data from other apps.
2. Visualization. Shiny app
3. Analysis. See associations between data types.
4. Experimentation framework. Run experiments on yourself

```

2. Briefly describe the background if that's not obvious.  What field
   or area of interest is it related to?  (Just briefly, in a paragraph
   or two, maximum.) References, including web references, are welcome
   where available.

```q2

'''
This package design is inspired by the OpenFDA Developer Challenge
(Option 1, Level 1). A link to the developer Challenge announcement is
here

https://open.fda.gov/update/an-open-challenge-to-tap-public-data/. The
challenge is broken into three phases. The first phase is identifying
spikes based on drug queries and "bucketizing variables over time,
weight, gender or drug pairs." The project will address this level of
the challenge and will also provide tools for visualization and
analysis.
'''


Fitbits and other health tracking devices have taken off in the last few years, as data enthusiasts seek a finer granularity of insight, and also seek to personalize their data as part of movements such as the "Quantified Self" movement, which advocates "self knowledge through numbers". There has been talk of how Fitbits can be incorporated into medical care, and stories of how they have led users to look into conditions they wouldnt' have otherwise known they had.

However, a common complaint about the Fitbit is the lack of analytics provided for by the company; users are left to glance through a couple raw graphs of their numbers, and, failing to continue to learn more from their data, after the novelty wears off, Fitbits often end up stuffed away, unused, in a drawer.

This package will bring the Fitbit one step closer to what it can be by providing users with suggestions and the ability to do more advanced analysis of their health. Instead of single graph visualizations of raw numbers, we can allow people to view the impact of one part of their life on another by combining data. Finally, we hope to allow the Fitbit to remain an interactive tool for self analysis and tailoring of one's life to meet one's goals, and a source of continual learning, by enabling individuals to run experiments in their life, then easily view the impact of the treatment on their Fitbit data.


This package builds off the fitbitScraper (https://cran.r-project.org/web/packages/fitbitScraper/index.html) package.  We looked at what fitbit users might want in terms of analytics 
(https://community.fitbit.com/t5/Fitbit-com-Dashboard/dashboard-needs-better-analytics/m-p/1185213)

Other links
- don't think we can actually do anything about this person's complaints:
http://www.computerworld.com/article/2951751/personal-technology/why-i-ditched-my-fitbit-and-what-this-means-for-analytics.html
- Other sites that create fitbit analytics
  - https://rocketgraph.com/reports/16-fitbit-analytics, https://www.producthunt.com/posts/fitbit-analytics-2
  - https://datarunsdeep.com.au/blog/track-your-life-in-google-analytics/
```

3. Provide one or two use cases; that is, in a sentence or so say how
   you see someone using the package.

```q3

'''
Scientists will query the R package for adverse events and patient
information. The queries will allow them to pass arguments.

Scientists can also visualize the results of their queries.

Scientists will also be able to utilize the analytical functions to
discover assocations among variables.

Administrative organizations will query the R package for adverse
event patterns in order to monitor the drug market.  These queries can
also be parameterized by the user.
'''

Individuals can use the R package to match data from their personal health devices and apps together.

Individuals can use the R package and Shiny app to run experiments on their Fitbit data; for example, track when they consume alcohol, and visualize the result on their recorded Fitbit resting heartrate.

Individuals can run analytics on their data, such as an analysis of how long they usually spend between waking up, and leaving bed in the morning.

Individuals can visualize associations in their data, such as the result of exercise and increased step counts on their sleep at night.

Individuals can view fluctuations in their measures of health over time, such as how their heart rate corresponds to other events in their health as recorded by their other tracking devices.

Individuals can design a training program, or use a suggested one, and measure the effect on their weight over time.



```

4. If you can, list the main functions or other software you plan on
   including.  No need to implement anything now, but you should have
   an idea.

```q4
'''
The R package will provide the following functions: reading and
writing data, creating graphs and visualizations, running
parameterized queries against datasets, providing for download
capabilibities of both data and graphics.  The R package will also
provide functions for deeper analysis by the user.
'''

The R package will provide capability to read in Fitbit, Apple, and other personal health data, match data from multiple sources, visualize the data and associations between different types of data, perform analysis on the data, and run experiments over time by adding other types of data to their personal tracking data.

```

```{r, echo = FALSE, hide = TRUE}
none <- "none"
basic <- "basic"
proficient <- "proficient"
```

5. For languages C/C++, Java, Python, Ruby, please specify your
   proficiency: Proficient, basic knowledge, no knowledge.  (This
   helps us evaluate the proposal sometimes and offer suggestions.)

```{r q5}
languages <- list(
    ## replace the values below appropriately for each member
    radke = c(cxx = none,
               java = basic,
               python = proficient,
               ruby = none)
    ,
    layu = c(cxx = none,
               java = basic,
               python = basic,
               ruby = none)
)
```








