---
title: "Project Proposal"
output:
     html_document
---

__Proposal Deadline: Jan 30, 2017 11:59pm__


- Edit this markdown file by __replacing only__ what is there in the
chunks named `q0`, `q1`, `q2`, ..., `q5` with your team's appropriate
answers. These chunks have been filled in as an example.

- Place the filled in markdown in any one of the team member's private
directory with the same name `proposal.Rmd`. Do not change the name!


0. Who are the authors of this package? See format below.

```{r q0}
authors <- list(sunet1 = "Full Name",
                sunet2 = "Full Name")
```

1. What is your package trying to achieve?  Who is it aimed at?

```q1

1.  Querying.  The package will enable users to query the Adverse
Events Database and other databases on OpenFDA from R. For example,
they could query information about events and patients.

2.  Visualization.  The package will also provide tools to
visualize for an inquired drug the relationship between adverse event
counts/types and time, drug pair, patient's weight, age, gender and
other variables.

3.  Analysis.  It will also include three functions to analyze
if 1) there are spikes in adverse event happenings, 2) if spikes
are related to drug recalls or enforcement and then 3) other
visualizations such as principal component charts to see which
variables contribute the most to certain phenomena.

The package is aimed at analysts and scientists in pharmaceutical
companies, research institutions, health care organizations, and other
administrative organizations (i.e., FDA).

```

2. Briefly describe the background if that's not obvious.  What field
   or area of interest is it related to?  (Just briefly, in a paragraph
   or two, maximum.) References, including web references, are welcome
   where available.

```q2

This package design is inspired by the OpenFDA Developer Challenge
(Option 1, Level 1). A link to the developer Challenge announcement is
here

https://open.fda.gov/update/an-open-challenge-to-tap-public-data/. The
challenge is broken into three phases. The first phase is identifying
spikes based on drug queries and "bucketizing variables over time,
weight, gender or drug pairs." The project will address this level of
the challenge and will also provide tools for visualization and
analysis.

```

3. Provide one or two use cases; that is, in a sentence or so say how
   you see someone using the package.

```q3

Scientists will query the R package for adverse events and patient
information. The queries will allow them to pass arguments.

Scientists can also visualize the results of their queries.

Scientists will also be able to utilize the analytical functions to
discover assocations among variables.

Administrative organizations will query the R package for adverse
event patterns in order to monitor the drug market.  These queries can
also be parameterized by the user.

```

4. If you can, list the main functions or other software you plan on
   including.  No need to implement anything now, but you should have
   an idea.

```q4

The R package will provide the following functions: reading and
writing data, creating graphs and visualizations, running
parameterized queries against datasets, providing for download
capabilibities of both data and graphics.  The R package will also
provide functions for deeper analysis by the user.

```

```{r, echo = FALSE, hide = TRUE}
none <- "none"
basic <- "basic"
proficient <- "proficient"
```

5. For languages C/C++, Java, Python, Ruby, please specify your
   proficiency: Proficient, basic knowledge, no knowledge.  (This
   helps us evaluate the proposal sometimes and offer suggestions.)

```{r q5}
languages <- list(
    ## replace the values below appropriately for each member
    sunet1 = c(cxx = none,
               java = proficient,
               python = proficient,
               ruby = none)
    ,
    sunet2 = c(cxx = none,
               java = proficient,
               python = proficient,
               ruby = none)
)
```








